---
title: "Exploratory Analysis"
output: html_document
author: "Group Members: Savannah Umali-Jepson, Sarah Peng, Lufei Wang, Ming Yan "
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(ggplot2)
source("livingvsdeceaseddistribution.R")
source("waiting_organ_type.R")
source("donor_age_breakdown.R")
source("donor_trend.R")
```
## Introduction
### This is an exploration into organ donation data and includes data analysis on age distribution of donors, how deceased vs living donor counts have changed over time, and the relationship between the donation waitlist, organ type, and blood type. We got our data from [Table of Contents for Results Tables for the 2019 National Survey of Organ Donation Attitudes and Practices (NSODAP)](https://www.organdonor.gov/about-dot/grants/research-reports) and [Donors Recovered in the U.S. by Donor Type](https://optn.transplant.hrsa.gov/data/view-data-reports/national-data/#) .

## Distribution of Variables

```{r, echo=FALSE}
mean_ages_plot

```

### This graph displays the distribution of the average number of donors per year by age group from 1988-2020. From this graph we can see that adults (ages 18-64), donate more than any other age group. When we think about this, it makes sense since most of the population is adults. While the elderly (ages 65+) are most likely to have more motivation to donate their organs than other age groups, since they likely have a shorter remaining life expectancy, we can see that older adults (ages 50-64) make up a larger percent of donors than elderly adults. Why would this be the case? While the elderly may be more willing, they are not likely to have as healthy organs, due to their age, making them less viable candidates to donate. Children (ages 0-17) on average donate the least. This is because children have the largest remaining life expectancy, than any other group. Most parents would not willingly allow their children to donate unless they are deceased or there are extreme circumstances. Overall this data follows a standard bell curve and has expected and understandable results. 


```{r, echo=FALSE}
plot2

```

### In this graph, we see the total of donors from 1988 to 2020 in 9 age groups. The bar graph shows the range from 1988-2020 throughout each age group. From the graph, we see that the most organ donations came from people in the age range of 50-64 years old. The age groups with the fewer donations besides the Unknown, are <1 and 1-5 years old.

## Relationships between variables

### Ming 
```{r, echo=FALSE}
donor_trend
    
```

### Throughout the last 20 years, there has been a growing trend for both types of the donors: deceased and living. Though there is a downward trend from 2005 to 2015 for the living donors, the total increase in donations is more than double comparing to the number of donations in 1998. The number of deceased donors has been gradually increasing, meaning that more and more people are willing to donate once they die. Over the years, the number of deceased donations has almost doubled, and almost every year there have been more deceased donors than living. While it is not surprising that there are more deceased donors than living, the rate at which the deceased donations are growing is surprising. While more people throughout all age-groups are familiar and willing to donate their organs, survey result shows that only 10% of them who would do it after their deaths. This would indicate a slower deceased growth rate than what we are seeing. A deeper investigation with more data would be conducted in the future to verify results.

```{r, echo=FALSE}
waiting_plot
```

### This graph displays the relationship between organ type and the number of people waiting for that organ by blood type. Top 5 in-demand organ was chosen to display. My independent variable is continous which is the number of people wahiting for a organ, and my dependent variable is categorical which is the type of organ a person is waiting for. I chose look at the numbers by blood type. From the graph, I found that kidney is the most in-demand organ. The demand for kidney is about 8 times as high as the second in-demand organ. The reason I believe is that although kidney can come from living donors, kidney failure is so rampant that there are still not enough donated organs to meet the demand. 
### There are more people with blood type O waiting for donation for all organs. I believe that is because people blood type O are compatible only with organs donated by another type O person. However, O blood type is universal donors because all other blood types are also compatible with type O blood. Therefore, it is harder for type O blood person to receive the right organ. Perple with AB blood type. in contrast, has the less amount of people waiting for donation because AB blood type is compatible with blood types A,B,AB or O and they are the universal recipients. 